---
output: rmarkdown::github_document
---

`epidata` : Tools to Retrieve Economic Policy Institute Data Library Extracts

The [Economic Policy Institute](http://www.epi.org/data/) provides researchers, media, and
the public with easily accessible, up-to-date, and comprehensive historical data on the 
American labor force. It is compiled from Economic Policy Institute analysis of government
data sources. Use it to research wages, inequality, and other economic indicators over 
time and among demographic groups. Data is usually updated monthly.

The following functions are implemented:

- `get_black_white_wage_gap`:	Retreive the percent by which hourly wages of black workers 
   are less than hourly wages of white workers
- `get_college_wage_premium`:	Retreive the percent by which hourly wages of college graduates 
   exceed those of otherwise equivalent high school graduates
- `get_employment_to_population_ratio`:	Retreive the share of the civilian noninstitutional 
   population that is employed
- `get_gender_wage_gap`:	Retreive the percent by which hourly wages of female workers are 
   less than hourly wages of male workers
- `get_hispanic_white_wage_gap`:	Retreive the percent by which hourly wages of Hispanic 
   workers are less than hourly wages of white workers
- `get_labor_force_participation_rate`:	Retreive the share of the civilian noninstitutional
   population that is in the labor force
- `get_long_term_unemployment`:	Retreive the share of the labor force that has been unemployed 
   for six months or longer
- `get_median_and_mean_wages`:	Retreive the hourly wage in the middle of the wage distribution
- `get_non_high_school_wage_penalty`:	Retreive the percent by which hourly wages of workers 
   without a high school diploma (or equivalent) are less than wages of otherwise equivalent workers who have graduated from high school
- `get_underemployment`:	Retreive the share of the labor force that is "underemployed"
- `get_unemployment`:	Retreive the share of the labor force without a job
- `get_unemployment_by_state`:	Retreive the share of the labor force without a job (by state)
- `get_wages_by_education`:	Retreive the average hourly wages of workers disaggregated by the 
   highest level of education attained
- `get_wages_by_percentile`:	Retreive wages at ten distinct points in the wage distribution
- `get_wage_ratios`:	Retreive the level of inequality within the hourly wage distribution.

### Installation

```{r eval=FALSE}
devtools::install_git("https://gitlab.com/hrbrmstr/epidata.git")
```

```{r message=FALSE, warning=FALSE, error=FALSE}
options(width=120)
```

### Usage

```{r message=FALSE, warning=FALSE, error=FALSE}
library(epidata)

# current verison
packageVersion("epidata")

get_black_white_wage_gap()

get_underemployment()

get_median_and_mean_wages("gr")
```

### Test Results

```{r message=FALSE, warning=FALSE, error=FALSE}
library(epidata)
library(testthat)

date()

test_dir("tests/")
```

